---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Downloads and unzips the data before storing it in the "data" dataframe
```{r}
##getting the file from its url
url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, "repdata-data-activity.zip", mode="wb")
unzip("repdata-data-activity.zip")
data <- read.csv("activity.csv")
##preprocessing the file

```
## What is mean total number of steps taken per day?
Agregates the number of steps taken each date and plots in an histogram
Calculates the mean an median values of steps per day
```{r}
##aggregating the number of steps for each day
stepsbyday <- aggregate(data$steps, by=list(data$date), na.rm=TRUE, FUN=sum)
names(stepsbyday)<-c("day","steps")
##generating the histogram
hist(stepsbyday$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
##calculating the mean and median values
mean_steps <- mean(stepsbyday$steps)
median_steps <- median(stepsbyday$steps)
```

The mean total number of steps by day is `r mean_steps`.
The median total number of steps by day is `r median_steps`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
