---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Download and unzip the data before storing it in the "data" dataframe
```{r}
##getting the file from its url
url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, "repdata-data-activity.zip", mode="wb")
unzip("repdata-data-activity.zip")
data <- read.csv("activity.csv")
##preprocessing the file

```
## What is mean total number of steps taken per day?
Agregate the number of steps taken each date and plot in an histogram
```{r}
##aggregating the number of steps for each day
stepsbyday <- aggregate(data$steps, by=list(data$date), FUN=sum)
names(stepsbyday)<-c("day","steps")
##generating the histogram
hist(stepsbyday$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
